### 2023-09-22 main.tex
fixed

@@

@@

 We assume that the boundary conditions of the above difference differential system are given and denoted by $\rho_{a}^{L}$ and $\rho_{a}^{R}$ for all $a\in\{1,\ldots,N\}$ at left and right respectively, further assuming that $\sum_{a=1}^{N}\rho_{a}^{L}=\nu$.
In the following of this section we answer to these questions. We first define the process on a chain in order to obtain \eqref{NSpeciesRD} for the evolution of the average occupation variable of type $a$. Once we have found such a process, we construct the Lie algebraic description and we use it to prove absorbing duality in this general case. 

@@

	\mathcal{L}^{rd}=+\mathcal{L}_{right}^{rd}\sum_{x=1}^{L-1}\mathcal{L}_{x,x+1}^{rd}+\mathcal{L}_{right}^{rd}

@@

defined $\forall c\in\{1,\ldots,N-2\}$ as 

@@

		1\quad &\text{if}\quad k=N

@@

		\alpha_{a}&=\sigma_{11}\rho_{a}^{L}+\sigma_{12}\sum_{b=2\,:\,b\neq a}^{N}\rho_{b}^{L}\qquad \text{if} \quad a\in \{1,\ldots,N-1\}\\

@@

		\sum_{a=2}^{N}n_{a}^{x}n_{q}^{x+1}-\left(\nu-\sum_{i=2}^{N}n_{i}^{x}\right)n_{q}^{x+1}-\sum_{\ell=2}^{N}n_{d}^{x}n_{\ell}^{x+1}-n_{d}^{x}\left(\nu-\sum_{i=2}^{N}n_{i}^{x+1}\right)

@@

where we call $q$ such that $g_{i}^{-1}(d)=q$. Let us observe that, by the definition of the map $g_{c}(\cdot)$ and for the fact that $1\leq c\leq N-2$, we have that $q\neq d$ and $q\neq 1$ .

@@

		\nu\sigma_{11}\rho_{d}^{L}+\nu\sigma_{12}\left(\sum_{b=2\,:\,b\neq d}^{N}\rho_{b}^{1}\right)-n_{d}^{1}\nu(\sigma_{11}+(N-2)\sigma_{12})
Using equation \eqref{actionGraphRD} we have that $\forall x\in \{1,\ldots,L-1\}$ and $\forall d\in\{1,\ldots,N-1\}$

@@

	\mathcal{L}^{rd}n_{d}^{1}=\sigma_{11}\nu \left(\rho_{d}^{1}-2n_{d}^{1}+n_{2}^{2}\right)+\sigma_{12}\nu\sum_{b=1\,:\,b\neq d}^{N-1}\left(\rho_{c}^{1}-2n_{c}^{1}+n_{c}^{2}\right)+\Upsilon\sum_{b=1\,:\,b\neq d}^{N-1}\left(n_{c}^{1}-n_{d}^{1}\right)

@@

where $\mathbb{E}_{\mu}$ denotes the expectation with respect to the law of the Markov process $(\bm{n})_{t\geq 0}$ initialized with arbitrary initial measure $\mu$. These average occupations evolve according to the Master equation

@@

Therefore, using the action of $\mathcal{L}^{rd}$ expressed by \eqref{DifferenceEquation} and \eqref{actionBDLine} we obtain a system of $N-1$ difference-differential equation where the discrete Laplacians of all species are coupled and linear reaction terms are present. For these reasons, this model is a generalization to $N$ species and to maximal occupation $\nu$ of the one studied in \cite{casini2022uphill}. This answers the first question of the beginning of this section.

@@

This can be proved just by imposing the detailed balance conditions for the edge $x,x+1$ and for the boundary generators. For fixed edge $x\in \{1,\ldots,L\}$ we only prove the detailed balance for  $\mathcal{L}_{x,x+1}^{c}$ with $c\in\{1,\ldots,N-2\}$ since for the other generators the computation is similar. Consider the transition $\bm{n}\to \bm{n}-\delta_{a}^{x}+\delta_{g_{c}(b)}^{x}+\delta_{g_{c}(b)}^{x+1}-\delta_{b}^{x+1}$ for arbitrary $a,b,\in\{1,\ldots,N-1\}$; the detailled balance reads
		&n_{a}^{x}n_{b}^{x+1}\frac{\nu!}{n_{1}^{x}!\cdots n_{a}^{x}!\cdots n_{g_{c}(b)}^{x}!\ldots n_{N}^{x}!}\frac{\nu!}{n_{1}^{x+1}!\cdots n_{b}^{x+1}!\cdots n_{g_{c}(a)}^{x+1}!\ldots n_{N}^{x+1}!}\\&\rho_{1}^{n_{1}^{x}+n_{1}^{x+1}}\cdots \rho_{a}^{n_{a}^{x}}\cdots\rho_{g_{c}(b)}^{n_{g_{c}(b)}^{x}}\cdots\rho_{b}^{n_{b}^{x+1}}\cdots \rho_{g_{c}(a)}^{n_{g_{c}(a)}^{x+1}}

@@

		\rho_{1}^{n_{1}^{x}+n_{1}^{x+1}}\ldots \rho_{a}^{n_{a}^{x}-1}\cdots\rho_{g_{c}(b)}^{n_{g_{c}(b)+1}^{x}}\cdots\rho_{b}^{n_{b}^{x+1}-1}\cdots \rho_{g_{c}(a)}^{n_{g_{c}(a)-1}^{x+1}}

@@

	\rho_{a}\rho_{b}=\rho_{g_{c}(b)}\rho_{g_{c}(a)}
that is true if and only if $\rho_{a}=\rho$ $\forall a\in \{1,\ldots,N-1\}$.\\
By similar computation it is possible to show that the same claim is obtained imposing the detailled balance for the transition $\bm{n}\to \bm{n}-\delta_{a}^{x}+\delta_{b}^{x}$, related to the generator $\mathcal{L}^{m}_{x,x+1}$. \\

@@

To sites $1,L$ we associate an ``extra site'', denoted by $0$ and $L+1$ respectively,

@@

We write the configurations $\bm{\xi} \in \widetilde\Omega$  as

@@

and the component $\xi_{a}^{0}$  gives the number of dual particles of type $a\in \{1,\ldots,N\}$ at 
the extra-site $0$ connected to $1$ and similarly for $xi_{a}^{L+1}$. \\

@@

$\mathcal{L}_{x,x+1}^{rd}$ is defined in \eqref{edgeGeneratorRD} and, $\widetilde{\mathcal{L}}_{left}^{rd}$ and $\widetilde{\mathcal{L}}_{right}^{rd}$ are again the absorbing dual generators $\widetilde{\mathcal{L}}_{x}$ defined in \eqref{siteDualGenerator} and specialized for sites $x=1$ and $x=L$. 
Observe that the generator $\widetilde{\mathcal{L}}^{rd}$ has the same dynamics as $\mathcal{L}^{rd}$ in the bulk, while the boundaries are absorbing. 

@@

The duality between the boundary generators is trivial, because they have the same form of \eqref{siteDualGenerator}, up to a specific choice of the parameters. Therefore we only need to prove the duality relation for the bulk generator. 
In the spirit of what we did in Section \ref{sectionDuality}, it is convenient to write the Hamiltonian of the process, that is defined through the basis elements \eqref{actionE} of the $gl(N)$ Lie algebra as 

@@

That is equivalent of showing 

@@

Since the matrix $R_{x}$ is the same, we shall show that $\mathcal{H}_{x,x+1}^{c}$ and $\mathcal{H}_{x,x+1}^{m}$ have $\exp{(E^{x})}\exp{(E^{x})}$ as a symmetry. The first result will follow from the commutator between $\mathcal{H}_{x,x+1}^{c}$ and $\sum_{d=1}^{N-1}(E_{dN}^{x})\otimes \sum_{k=1}^{N-1}(E_{bN}^{x})$. Using \eqref{eq:comgl}, the bilinearity and associativity of the Kronecker product and the bilinearity of the brackets we obtain 

@@

The second result will follow from the commutator between $\mathcal{H}_{x,x+1}^{m}$ and $\sum_{d=1}^{N-1}(E_{dN}^{x})\otimes \sum_{k=1}^{N-1}E_{bN}^{x}$. The prof is similar and gives

@@

\appendix
\begin{comment}
{\color{blue}
\section{Oscillator realization}\label{appA}

>>>>

 A new technique to compute absorption probabilities has been proposed in \cite{floreani2023non}.

@@

\begin{equation}\label{abs-dual-boundary-chain}

@@

 We assume that the boundary conditions of the above difference differential system are given and denoted by $\rho_{a}^{0}$ and $\rho_{a}^{L+1}$ for all $a\in\{1,\ldots,N-1\}$ at left and right respectively, further assuming that $\sum_{a=1}^{N}\rho_{a}^{0}=\nu$.
 We first define the process on a chain in order to obtain \eqref{NSpeciesRD} for the evolution of the average occupation variable of type $a$. Once we have found such a process, we construct the Lie algebraic description and we use it to prove absorbing duality.

@@

	\mathcal{L}^{rd}=\mathcal{L}_{right}^{rd}\sum_{x=1}^{L-1}\mathcal{L}_{x,x+1}^{rd}+\mathcal{L}_{right}^{rd}

@@

defined as

@@

		N\quad &\text{if}\quad k=N

@@

		\alpha_{a}&=\sigma_{11}\rho_{a}^{0}+\sigma_{12}\sum_{b=1\,:\,b\neq a}^{N-1}\rho_{b}^{0}\qquad \text{if} \quad a\in \{1,\ldots,N-1\}\\

@@

		\sum_{a=1}^{N-1}n_{a}^{x}n_{q}^{x+1}+\left(\nu-\sum_{i=1}^{N-1}n_{i}^{x}\right)n_{q}^{x+1}-\sum_{\ell=1}^{N-1}n_{d}^{x}n_{\ell}^{x+1}-n_{d}^{x}\left(\nu-\sum_{i=1}^{N-1}n_{i}^{x+1}\right)

@@

where we call $q$ such that $g_{i}^{-1}(d)=q$ and we used the fact that $n_{N}^{x}=\nu-\sum_{i=1}^{N-1}n_{i}^{x}$. Let us observe that, by the definition of the map $g_{c}(\cdot)$ and for the fact that $1\leq c\leq N-2$, we have that $q\neq d$ and $q\neq N$ .

@@

		\nu\sigma_{11}\rho_{d}^{0}+\nu\sigma_{12}\left(\sum_{b=1\,:\,b\neq d}^{N-1}\rho_{b}^{0}\right)-n_{d}^{1}\nu(\sigma_{11}+(N-2)\sigma_{12})
		\nu \sigma_{11} (\rho_{d}^{0}-n_{d}^{1})+\nu\sigma_{12}\sum_{b=1\,:\,b\neq d}^{N-1}(\rho_{b}^{0}-n_{d}^{1})
Using equation \eqref{actionGraphRD} we have that $\forall x\in \{2,\ldots,L-1\}$ and $\forall d\in\{1,\ldots,N-1\}$

@@

	\mathcal{L}^{rd}n_{d}^{1}=\sigma_{11}\nu \left(\rho_{d}^{0}-2n_{d}^{1}+n_{d}^{2}\right)+\sigma_{12}\nu\sum_{b=1\,:\,b\neq d}^{N-1}\left(\rho_{b}^{0}-2n_{b}^{1}+n_{b}^{2}\right)+\Upsilon\sum_{b=1\,:\,b\neq d}^{N-1}\left(n_{b}^{1}-n_{d}^{1}\right)

@@

where $\mathbb{E}_{\mu}$ denotes the expectation with respect to the law of the Markov process $(\bm{n}(t))_{t\geq 0}$ initialized with arbitrary initial measure $\mu$. These average occupations evolve according to the Master equation

@@

Therefore, using the action of $\mathcal{L}^{rd}$ expressed by \eqref{DifferenceEquation} and \eqref{actionBDLine} we obtain a system of $N-1$ difference-differential equation where the discrete Laplacians of all species are coupled and linear reaction terms are present. For these reasons, this model is a generalization to $N$ species and to maximal occupation $\nu$ of the one studied in \cite{casini2022uphill}. This answers the first question we addressed to in the beginning of this section.

@@

This can be proved just by imposing the detailed balance conditions for the edge $x,x+1$ and for the boundary generators. For fixed site $x\in \{1,\ldots,L\}$ we only prove the detailed balance for  $\mathcal{L}_{x,x+1}^{c}$ with $c\in\{1,\ldots,N-2\}$ since for the other generators  $\mathcal{L}_{x,x+1}^{m}$, $\mathcal{L}_{left}^{rd}$ and $\mathcal{L}_{right}^{rd}$ the computation is similar. Consider the transition $\bm{n}\to \bm{n}-\delta_{a}^{x}+\delta_{g_{c}(b)}^{x}+\delta_{g_{c}(b)}^{x+1}-\delta_{b}^{x+1}$ for arbitrary $a,b\in\{1,\ldots,N-1\}$ and $c\in\{1,\ldots,N-2\}$; the detailed balance reads
		&n_{a}^{x}n_{b}^{x+1}\frac{\nu!}{n_{1}^{x}!\cdots n_{a}^{x}!\cdots n_{g_{c}(b)}^{x}!\ldots n_{N}^{x}!}\frac{\nu!}{n_{1}^{x+1}!\cdots n_{b}^{x+1}!\cdots n_{g_{c}(a)}^{x+1}!\ldots n_{N}^{x+1}!}\\&\rho_{1}^{n_{1}^{x}+n_{1}^{x+1}}\cdots \rho_{a}^{n_{a}^{x}}\cdots\rho_{g_{c}(b)}^{n_{g_{c}(b)}^{x}}\cdots\rho_{b}^{n_{b}^{x+1}}\cdots \rho_{g_{c}(a)}^{n_{g_{c}(a)}^{x+1}}\cdots \rho_{N}^{n_{N}^{x}+n_{N}^{x+1}}

@@

		\rho_{1}^{n_{1}^{x}+n_{1}^{x+1}}\ldots \rho_{a}^{n_{a}^{x}-1}\cdots\rho_{g_{c}(b)}^{n_{g_{c}(b)}^{x}+1}\cdots\rho_{b}^{n_{b}^{x+1}-1}\cdots \rho_{g_{c}(a)}^{n_{g_{c}(a)}^{x+1}+1}

@@

	\rho_{a}\rho_{b}=\rho_{g_{c}(a)}\rho_{g_{c}(b)}
By the arbitrariness of $a,b,$ it is true if and only if $\rho_{a}=\rho$ $\forall a\in \{1,\ldots,N-1\}$.\\

@@

We associate an ``extra site'', denoted by $0$ and $L+1$ attached to $1$ and $L$ respectively,

@@

We denote the configurations $\bm{\xi} \in \widetilde\Omega$  as

@@

and the component $\xi_{a}^{0},\xi_{a}^{L+1}$  give the number at the extra sites. \\

@@

$\mathcal{L}_{x,x+1}^{rd}$ is defined in \eqref{edgeGeneratorRD} and, $\widetilde{\mathcal{L}}_{left}^{rd}$ and $\widetilde{\mathcal{L}}_{right}^{rd}$ are again the absorbing dual generators $\widetilde{\mathcal{L}}_{x}$ defined in \eqref{abs-dual-boundary-chain}, with the choice of the reservoirs parameters \eqref{boundaryParamRD}.

@@

We only need to prove duality for the bulk generator, since the boundary generators are the same as in Subsection \ref{Subsection-ss-nonI}. 
In the spirit of Section \ref{sectionDuality}, it is convenient to write the Hamiltonian of the process, that is defined through the basis elements \eqref{actionE} of the $gl(N)$ Lie algebra as 

@@

It is enough of showing 

@@

Since the matrix $R_{x}$ is the same as in Section \ref{sectionDuality}, we shall show that $\mathcal{H}_{x,x+1}^{c}$ and $\mathcal{H}_{x,x+1}^{m}$ have $\exp{(E^{x})}\exp{(E^{x+1})}$ as a symmetry.

\paragraph{Duality for $H_{x,x+1}^{c}$} We compute the commutator between $\mathcal{H}_{x,x+1}^{c}$ and $\sum_{d=1}^{N-1}(E_{dN}^{x})\otimes \sum_{k=1}^{N-1}(E_{bN}^{x})$. Using \eqref{eq:comgl}, the bilinearity and associativity of the Kronecker product and the bilinearity of the brackets we obtain 

@@

\paragraph{Duality for $H_{x,x+1}^{m}$} We compute the commutator between $\mathcal{H}_{x,x+1}^{m}$ and $\sum_{d=1}^{N-1}(E_{dN}^{x})\otimes \sum_{k=1}^{N-1}E_{bN}^{x}$. The prof is similar and gives

@@

\appendix
\section*{Appendix}
\begin{comment}
{\color{blue}
\section{Oscillator realization}\label{appA}
